---
title: "regression_summ_stats"
format: html
editor: visual
---

Makes summary statistics for regression variables in trip_catch_2022_piv_acs.RData, from making_trip_catch_2022_piv_acs.RData. For more see izzy_zheng_readme.txt.

setup

```{r}
#| warning: false
library(tidyverse)
library(here)
library(psych)
data_path <- here('data')
load(here('data','intermediate_datasets',"trip_catch_2022_piv_acs.RData"))
percent_eaten_reg_data <- trip_catch_2022_piv_acs |>
  filter(tot_cat_1 >0)|>
  select(unemployment,income,not_english,black,nonwhite,
         noncit, pct_pov, bad_english, college, white, 
         indigenous, asian, pac_islander, hispanic,
         mode_reg,hrsf,age,ffdays2,cntrbtrs,gender,wave,
         intsite,percent_eaten, st_res,wp_int,desired_fish)|>
  na.omit()
```

divide dataset into controls, intermediate variables, and dependent + independent variables, filter out character variables

```{r}
str_vars <- percent_eaten_reg_data |>
  select(where(is.character),gender)
#intermed_vars <- percent_eaten_reg_data |>
  #select(ID_CODE,starts_with('species'),starts_with('claim'),
         #starts_with('harvest'),starts_with('landing'),starts_with('release'),
         #starts_with('wgt'),starts_with('tot_cat'),wp_int)|>
  #select(where(is.numeric))
dep_indep <-percent_eaten_reg_data |>
  select(percent_eaten,#dep var
         unemployment,pct_pov,income,college, #richness
         noncit,not_english,bad_english,#immigrant
         black,nonwhite,white,asian,#race/eth
         indigenous,pac_islander,hispanic,nonwhite #race/eth
         )|>
  select(where(is.numeric))
controls <- percent_eaten_reg_data |>
  select(#-(intermed_vars|>names()),
         -(dep_indep|>names()))|>
  select(where(is.numeric))
         

```

summary statistics - controls (may be categorical or not)

```{r}
controls_summstat <- describe(controls,fast=TRUE)
#intermed_summstat <- describe(intermed_vars,fast=TRUE)
depindep_summstat <- describe(dep_indep,fast=TRUE)
```

```{r}
controls_summstat|>select(-vars,-range,-se,-n)
#intermed_summstat|>select(-vars,-range,-se,-n)
depindep_summstat|>select(-vars,-range,-se,-n)
```

summary statistics - frequency tables of categorical variables (may include controls)

```{r}
kod_freq <- table(str_vars$kod)|>as.data.frame()
prim1_common_freq <- table(str_vars$prim1_common)|>as.data.frame()
prim2_common_freq <- table(str_vars$prim2_common)|>as.data.frame()
common_fish_freq <- table(trip_catch_2022_piv_acs$common_fish)|>as.data.frame()
```

```{r}
kod_freq
prim1_common_freq
prim2_common_freq
common_fish_freq
```

```{r}
zip_freq <- table(str_vars$zip)|>as.data.frame()
county_freq <- table(str_vars$county)|>as.data.frame()
date1_freq <- table(str_vars$date1)|>as.data.frame()
```

```{r}
zip_freq
county_freq
date1_freq
```

```{r}
gender_freq <- table(str_vars$gender)|>as.data.frame()
f_by_p_freq <- table(str_vars$f_by_p)|>as.data.frame()
gear_freq <- table(str_vars$gear)|>as.data.frame()
intsite_freq <- table(str_vars$intsite)|>as.data.frame()
```

```{r}
gender_freq 
f_by_p_freq 
gear_freq
intsite_freq
```

```{r}
leader_freq <- table(str_vars$leader)|>as.data.frame() #honestly probably don't need to use this variable
species_1_freq <- table(str_vars$species_1)|>as.data.frame()
species_2_freq <- table(str_vars$species_2)|>as.data.frame()
species_3_freq <- table(str_vars$species_3)|>as.data.frame()
species_4_freq <- table(str_vars$species_4)|>as.data.frame()
species_5_freq <- table(str_vars$species_5)|>as.data.frame()
species_6_freq <- table(str_vars$species_6)|>as.data.frame()
species_7_freq <- table(str_vars$species_7)|>as.data.frame()
#would be good to wrangle a new variable that is species they caught the most of in weight
```

```{r}
leader_freq 
species_1_freq 
species_2_freq 
species_3_freq 
species_4_freq 
species_5_freq 
species_6_freq 
species_7_freq 
```

```{r}
st_res_freq<-table(percent_eaten_reg_data$st_res)
st_res_freq
```
