---
title: "2022_visualizations"
format: html
editor: visual
---

## 

Setup

```{r}
library(tidyverse)
library(here)
library(readr)
library(sf)
library(mapview)
library(RColorBrewer)
```

```{r}
#load data
data_path <- here('data') #relative path to the data folder
```

load data. These datasets were made in the making_intermediate_datasets file in "Programs".

```{r}
load(here('data','intermediate_datasets','cb_zip.RData'))
load(here('data','intermediate_datasets','manual_zip.RData'))
load(here('data','intermediate_datasets','trip_catch_2022.RData'))
```

Visualizations

group by zip code

```{r}
tripcatch_2022_zip <- trip_catch_2022 |>
  group_by(ZIP)|>
  reframe(num_eaten = sum(landing),
          num_angler_trips = ID_CODE|>length())|>
  mutate(fish_per_angler_trip=num_eaten/num_angler_trips)
#represents fish per person per trip
```

join zip and tripcatch together

```{r}
tripcatch_2022_zip <- zip_manual |> 
  left_join(tripcatch_2022_zip,by=c("ZIP"="ZIP"))

#if we do a right join with zip, you'll get all the zip codes but no relative knowledge of where they are
```

view map

```{r}
mapview(tripcatch_2022_zip,
      zcol = 'num_eaten',
          col.regions=brewer.pal(9, "Blues"),
          alpha.regions = 0.8,
          layer.name = '# Fish Eaten')
```

is the variation from one superfisher or a sizable community of fishers? i think this codeblock is useless but keeping it

```{r}
#get vector of cb zip codes w 20+ fish/angler
many_fish_zip <- tripcatch_2022_zip_joined |>
  filter(fish_per_angler > 20)|>
  st_drop_geometry()|>
  select(ZIP)

#join to catch_2022 to look at harvest per fisher
many_fish_zip <- many_fish_zip |>
  left_join(tripcatch_2022_zip,by=c("ZIP"="ZIP"))

#visualization: ideally some visualization of number of fishermen in each zip code. x axis num fisherman, y axis number of zip codes
ggplot(data = many_fish_zip, aes(x=num_anglers))+geom_histogram()
#shows that the zip codes w 20+ fish/trip have 1 or 2 trips in them; so super dark areas are not nec places where lots of fish consumption is happening
#a better indicator might be total fish/population, or total fish, or filtering to areas that have at least 3 anglers in them
#avg num anglers=7, median anglers = 3
```

```{r}
#more data visualizations
ggplot(data = tripcatch_2022_zip, aes(x=fish_per_angler_trip))+geom_histogram()
#shows HEAVY skew right, most common value of fish/angler is 0 at the zipcode level
#max number of fish/angler is about 32
```

map of number of trips

```{r}
mapview(tripcatch_2022_zip,
      zcol = 'num_anglers',
          col.regions=brewer.pal(9, "Blues"),
          alpha.regions = 0.8,
          layer.name = 'Per Angler Per Trip Frequency')
```

Amt of fish by weight

```{r}
#sum the weight of each zip code's trip's total fish
weight_tripcatch <- trip_catch_2022 |>
  group_by(ID_CODE)|>
  summarize(weight_claim = sum(wgt_a),
    weight_per_angler_per_trip = sum(wgt_a)/sum(CNTRBTRS))#should be 5444 obs
#note that this is by claim, not landing
#if we wanted to do landing, we'd have to edit the #cntrbtrs so that it shares values with the LEADER variable

#aggregate at zip code level
weight_tripcatch_zip <- trip_catch_2022 |>
  group_by(ID_CODE,ZIP)|>
  summarize()|>
  inner_join(weight_tripcatch,by=c("ID_CODE"="ID_CODE"))|>#to add ZIP variable
  group_by(ZIP)|>
  reframe(zip_weight_per_angler_trip = sum(weight_per_angler_per_trip),
          zip_weight = sum(weight_claim))
#join to zip
weight_tripcatch_zip <- zip_manual |>
  left_join(weight_tripcatch_zip,by=c("ZIP"="ZIP"))
#map it
  mapview(weight_tripcatch_zip,
        zcol="zip_weight",
        col.regions=brewer.pal(9, "Blues"),
          alpha.regions = 0.8,
          layer.name = 'Fish Weight') 
#add the other zip_manual shapefile
```

Amt of fish in weight per angler

```{r}
mapview(weight_tripcatch_zip,
        zcol="zip_weight_per_angler_trip",
        col.regions=brewer.pal(9, "Blues"),
          alpha.regions = 0.8,
          layer.name = 'Weight per angler')
```

Bivariate mapping (ACS data): show relationship between race/income + fish consumption

```{r}

```

bivariate mapping data wrangling: join acs data in block group format to trip catch in zip format

```{r}
#clean acs data for joining (thanks peiley!)
cb_acs_df_clean <- cb_acs_df %>%
    st_drop_geometry() %>%
    pivot_wider(names_from=variable,values_from=estimate) %>%
    mutate(
           #compute percent below poverty threshold -->
      # pov50 (# below 0.5) and pov 99 (# between 0.5 and 0.99)
           pov99=(pov99+pov50)/pop*100,
           pov50=pov50/pop*100,
           income=income,
           state = substr(GEOID, 1, 2))

#convert acs to zipcode level using bg_zip
zip_acs <- cb_acs_df_clean |>
  inner_join(bg_zip,by=c("GEOID"="GEOID_bg")) |>
  group_by(ZIP_final)|>
  reframe(pct_minority = (sum(black)+sum(indian)+sum(asian)+sum(hispanic))/sum(pop),
          pct_black = sum(black)/sum(pop),
          pct_50=sum(pov50)/sum(pop),#percent of ppl under half the fed poverty income threshold
          pct_100=(sum(pov50)+sum(pov99))/sum(pop),
          pct_indian = sum(indian)/sum(pop),
          pct_hispanic = sum(hispanic)/sum(pop),
          pct_asian=sum(asian)/sum(pop))#pct ppl under fed pov income threshold

#join with tripcatch by zip code
tripcatch_2022_acs <- tripcatch_2022_zip|>
  inner_join(zip_acs,by=c("ZIP"="ZIP_final"))
```

bivariate package data (biscale)

note: am confused about the breaks thing: equal vs quantile vs jenks vs fisher?

```{r}
library(biscale)
library(cowplot)
biv_data<-tripcatch_2022_acs |> select(-pct_minority,-pct_black)|>
  na.omit()
biv_data <- bi_class(biv_data, y=pct_100,x=fish_per_angler_trip,style="jenks",dim=3)
```

```{r}
map<-ggplot() +
  geom_sf(data=zip_manual,
          fill="White")+
  geom_sf(data=biv_data,
          mapping=aes(fill=bi_class),
          show.legend=FALSE)+
  bi_scale_fill(pal="Bluegill",dim=3)+
  bi_theme() 
legend<-bi_legend(pal="Bluegill",
                  dim=3,
                  ylab = "% Low Income",
                  xlab = "Fish per Angler")

final_plot <- ggdraw() +
  draw_plot(map,0,0,1,1)+
  draw_plot(legend,0.2,.65,.2,.2)
final_plot
```

same map but for % black

```{r}
black_data<-tripcatch_2022_acs |> select(-pct_minority,-pct_100)|>
  na.omit()
black_data <- bi_class(black_data, x=fish_per_angler,y=pct_black,style="jenks",dim=3)

map_black <-ggplot() +
  geom_sf(data=cb_zip,
          fill="White")+
  geom_sf(data=black_data,
          mapping=aes(fill=bi_class),
          show.legend=FALSE)+
  bi_scale_fill(pal="Bluegill",dim=3)+
  bi_theme() 
legend<-bi_legend(pal="Bluegill",
                  dim=3,
                  xlab = "Fish per Angler",
                  ylab = "% Black")

black_plot <- ggdraw() +
  draw_plot(map,0,0,1,1)+
  draw_plot(legend,0.2,.65,.2,.2)
black_plot
```

same map but with minorities

```{r}
minority_data<-tripcatch_2022_acs |> select(-pct_black,-pct_100)|>
  na.omit()
minority_data <- bi_class(minority_data, x=fish_per_angler,y=pct_minority,style="jenks",dim=3)

map_black <-ggplot() +
  geom_sf(data=cb_zip,
          fill="White")+
  geom_sf(data=minority_data,
          mapping=aes(fill=bi_class),
          show.legend=FALSE)+
  bi_scale_fill(pal="Bluegill",dim=3)+
  bi_theme() 
legend<-bi_legend(pal="Bluegill",
                  dim=3,
                  xlab = "Fish per Angler",
                  ylab = "% Minority")

minority_plot <- ggdraw() +
  draw_plot(map,0,0,1,1)+
  draw_plot(legend,0.2,.65,.2,.2)
minority_plot
```

ggplot ej groups against fish caught

```{r}
ggplot(tripcatch_2022_acs,aes(pct_100,fish_per_angler)) +
  geom_point() +
  labs(title="Fish per Angler vs % Under Poverty Threshold")
#shows vast majority of fishing zip codes get no fish and have no poor people
#loose negative correlation btwn fish/angler and percent under the pov threshold
```

ggplot with % black

```{r}
ggplot(tripcatch_2022_acs,aes(pct_black,fish_per_angler)) +
  geom_point() +
  labs(title="Fish per Angler vs % Black")
#shows vast majority of fishing zip codes get no fish and have some black people
#again, loose negative correlation btwn fish/angler and %black
```

```{r}
ggplot(tripcatch_2022_acs,aes(pct_minority,fish_per_angler)) +
  geom_point() +
  labs(title="Fish per Angler vs % Minority")
#shows vast majority of fishing zip codes get no fish and have decent percentage minorities
#again, loose negative correlation btwn fish/angler and % minority, but way less than before; implies nonblack minorities more likely to fish than blacks?
```
