---
title: "make_buffer_zip"
format: html
editor: visual
---

## 

This document will make an sf object that includes all zip codes within a 100 mile radius of the nearest sitepoint, called buffer_zip. Also creates active_sites, an sf that contains all active CB fishing sites. Also makes a map of all fishing zip codes that send fishers to the Chesapeake overlaid with the buffer_zip zips and the Chesapeake Bay watershed boundary, for context of what zip codes are being filtered into the dataset. For more see izzy_zheng_readme.txt.

Setup

```{r}
library(here)
library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
```

load data

```{r}
data_path <- here('data') #relative path to the data folder
zip <- st_read(here(data_path, 'zipcodes', 'tl_2020_us_zcta520.shp')) #zip code shapefiles originally taken from https://www2.census.gov/geo/tiger/TIGER2020/ZCTA520/
zip<- zip |> select(-CLASSFP20,-MTFCC20,-FUNCSTAT20,-ALAND20,-AWATER20)
access_df <- read.csv(here('data', 'NOAA MRF', 'Public Fishing Access Site Register','SITE_new.csv'))
cb_sf <- st_read(here(data_path, 'watershed boundary', 'Chesapeake_Bay_Watershed_Boundary.shp')) #Chesapeake Bay shapefile
load(here('data','intermediate_datasets','trip_catch_2022.RData'))
```

```{r}
cb_sf <- cb_sf|> st_transform(st_crs(zip)) 
```

filter sites to within the CB border.

```{r}
cb_access_sf <- access_df %>%
  rename_all(tolower) |>
  st_as_sf(coords = c(x = 'site_long',
             y = 'site_lat'),
             crs = 4326) |>
 st_transform(st_crs(zip)) |>
    st_filter(cb_sf) #filter to sites in the cheseapeake bay
save(cb_access_sf,file=here(data_path,"intermediate_datasets","active_sites.Rdata"))
```

add 100-mi buffers

```{r}
site_buffer <- st_buffer(cb_access_sf,dist = 160934.4) #meter conversion of 100 miles
#visualization:
tmap_leaflet(tm_shape(site_buffer) +
  tm_polygons(alpha=0)+
  tm_symbols(col = "red", size = 0.1))
```

filter to zip codes within buffers

```{r}
zip |>
  st_transform(st_crs(cb_sf)) 
buffer_zip <- st_filter(zip,site_buffer)
```

```{r}
save(buffer_zip,file=here(data_path,"intermediate_datasets","buffer_zip.Rdata"))
```

Visual of chesapeake bay, fishing zips

orange = CB boundary

blue=fishing zips

red = mapped zip codes

```{r}
fishing_zips <- trip_catch_2022 |>
  group_by(ZIP)|>
  reframe(fishing = TRUE)
fishing_zips <- zip |>
  inner_join(fishing_zips,by=c('ZCTA5CE20'='ZIP'))
tmap_leaflet(
  tm_shape(buffer_zip)+tm_polygons(col="red",alpha=0.8)+
  tm_shape(fishing_zips)+tm_polygons(col='blue',alpha=0.8)+
    tm_shape(cb_sf)+tm_polygons(border.col='orange',alpha=0))
```

```{r}

tmap_leaflet(
  tm_shape(buffer_zip)+tm_polygons(col="grey",alpha=0.8)+
  tm_shape(fishing_zips)+tm_polygons(col='blue',alpha=0.8)+
  tm_shape(site_buffer) + tm_symbols(col = "red", size = 0.1)) |> 
  setView(lat=38.5,lng=-77,zoom = 6.4) 
```
