---
title: "Untitled"
format: html
editor: visual
---

## tentative regressions

Run preliminary regressions on trip_catch_2022_piv_acs.RData, from making_trip_catch_2022_piv_acs.RData. For more see izzy_zheng_readme.txt.

```{r}
library(fixest)
library(here)
library(tidyverse)
data_path <- here('data')
load(here('data','intermediate_datasets',"trip_catch_2022_piv_acs.RData"))
```

proportion eaten

```{r}
#filter out NA's for percent_eaten; there are na's where tot_cat of the fisher == 0
percent_eaten_reg_data <- trip_catch_2022_piv_acs |>
  filter(tot_cat_1 >0)|>
  select(unemployment,not_english,hispanic,black,wp_int, income,nonwhite,asian,
         #mode_reg,hrsf,age,cntrbtrs,gender,wave,
         #site_id,percent_eaten, st_res,wp_int,
         dne_total, polluted_proportion,water_body,num_ramps,desired_fish,ffdays2,
         light_att, num_advisories
         )|>
  na.omit()

percent_eaten_reg_data <- percent_eaten_reg_data |> 
  mutate(catfish = (str_detect(toupper(species_1), "CATFISH")))
```

```{r}
dne_reg<-lm(data = percent_eaten_reg_data,formula= dne_total ~ unemployment + black + hispanic + log(income)+light_att + num_advisories,
      weights = percent_eaten_reg_data$wp_int)
dne_df<-summary(dne_reg)
dne_df
```

```{r}
prop_polluted_reg<-lm(formula= polluted_proportion ~ not_english + black + unemployment + log(income) + light_att,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
prop_polluted_df<-summary(prop_polluted_reg)
prop_polluted_df
```

```{r}
reg<-lm(formula= percent_eaten ~ unemployment + income + not_english + black + #ej vars
        mode_reg + hrsf + age + ffdays2 + cntrbtrs +
        gender + st_res,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
reg_df<-summary(reg)
reg_df
```

add intsite

```{r}
reg_site <-lm(formula = percent_eaten ~ income + unemployment + not_english + black + #ej vars
        mode_reg + hrsf + age + ffdays2 + cntrbtrs +
        gender + st_res + intsite, 
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
reg_site
```

add species

```{r}
reg_common_fish <-lm(formula = percent_eaten ~ income + unemployment + not_english + black + #ej vars
        mode_reg + hrsf + age + ffdays2 + cntrbtrs +
        gender + st_res + common_fish, 
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(reg_primcaught)
```

desired fish dummy

```{r}
reg_desired_fish <-lm(formula = percent_eaten ~ income + unemployment + not_english + black + #ej vars
        mode_reg + hrsf + age + ffdays2 + cntrbtrs +
        gender + st_res + desired_fish,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
```

```{r}
reg_nocharter <- lm(formula = percent_eaten ~ income + unemployment + not_english + black + #ej vars
        mode_reg + hrsf + age + ffdays2 + cntrbtrs +
        gender + st_res, 
  data=percent_eaten_reg_data |> filter(mode_reg != "with captain"),
      weights = percent_eaten_reg_data|> filter(mode_reg != "with captain")|>select(wp_int)|>pull())
```

```{r}
stargazer(bare_reg, reg,reg_site,reg_desired_fish, type="text",out=here('Results','RecFishing','RegTable4.txt'), title = "Regression Results")
```

move st res to fixed

```{r}
bare_reg<-feols(fml = percent_eaten ~ unemployment + income + not_english + black + 
                  wave + age+gender + ffdays2 + hrsf + cntrbtrs + prim1_common +
                  mode_reg +intsite + prim_caught  + st_res, #ej vars,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(bare_reg)
```

almost no controls version

```{r}
bare_reg<-lm(formula = percent_eaten ~ income + unemployment + not_english + black, #ej vars,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(bare_reg)
```

add wave

```{r}
bare_reg<-feols(fml = percent_eaten ~ unemployment + not_english + black + wave| st_res, #ej vars,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(bare_reg)
```

add age,gender

```{r}
bare_reg<-feols(fml = percent_eaten ~ unemployment + not_english + black + wave +
                  age+gender| st_res, #ej vars,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(bare_reg)
```

add ffdays2,hrsf

```{r}
bare_reg<-feols(fml = percent_eaten ~ unemployment + not_english + black + wave +
                  age+gender + ffdays2 + hrsf| st_res, #ej vars,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(bare_reg)
```

add cntrbtrs

```{r}
bare_reg<-feols(fml = percent_eaten ~ unemployment + not_english + black + wave +
                  age+gender + ffdays2 + hrsf + cntrbtrs| st_res, #ej vars,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(bare_reg)

```

add primm1 common

```{r}
bare_reg<-feols(fml = percent_eaten ~ unemployment + not_english + black + wave +
                  age+gender + ffdays2 + hrsf + cntrbtrs + prim1_common| st_res, #ej vars,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(bare_reg)
```

add mode

```{r}
bare_reg<-feols(fml = percent_eaten ~ unemployment + not_english + black + wave +
                  age+gender + ffdays2 + hrsf + cntrbtrs + prim1_common + mode_fx| st_res, #ej vars,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(bare_reg)
```

add intsite

```{r}
bare_reg<-feols(fml = percent_eaten ~ unemployment + not_english + black + wave +
                  age+gender + ffdays2 + hrsf + cntrbtrs + prim1_common + mode_fx
                +intsite| st_res, #ej vars,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(bare_reg)
```

add primcaught

```{r}
bare_reg<-feols(fml = percent_eaten ~ unemployment + not_english + black + wave +
                  age+gender + ffdays2 + hrsf + cntrbtrs + prim1_common + mode_fx
                +intsite + prim_caught| st_res, #ej vars,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(bare_reg)
```

add gear

```{r}
bare_reg<-feols(fml = percent_eaten ~ unemployment + not_english + black + wave +
                  age+gender + ffdays2 + hrsf + cntrbtrs + prim1_common + mode_fx
                +intsite + prim_caught  | st_res, #ej vars,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(bare_reg)
```

```{r}
bare_reg<-feols(fml = percent_eaten ~ unemployment + not_english + black #ej vars
                + wave + #month proxy
                  age+gender + 
                  ffdays2 + hrsf + #days fished in last 2 months, hoursefished
                cntrbtrs + prim1_common + #fish they were looking for
                  mode_fx +intsite + prim_caught   + st_res, #ej vars,
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(bare_reg)
```

why fbyp, change mode_f to categorical,

try clustering? like unemp + college + pov rate + income, noncit+not_english, modes clustered together, avidity clustered together, intsite by body of water/county, fish categories? see matt's old paper about categorizing fish

quartile ej version

```{r}
reg<-feols(fml = percent_eaten ~ unemployment_quartile + black_quartile + not_english_quartile + zip,#ej vars
  data=percent_eaten_reg_data,
      weights = percent_eaten_reg_data$wp_int)
summary(reg)
```
